{\itshape Advanced Robotics} A simple path optimization method for motion planning

### 2015-10-08 paper.tex
Replace transf constraints by orthonogal ones, re-introduce stack of jacobians

% Math shortcuts :
\newcommand\real{\mathbb{R}}
\newcommand\p{\mathbf{p}}
\newcommand\pii{\mathbf{p_{i,i+1}}}
\newcommand\pij{\mathbf{p_{i,j}}}
\newcommand\gT{\tilde{g}^T}
\newcommand\g{\tilde{g}}
\newcommand\CS{\mathcal{C}}
\newcommand\dimCS{n_\mathbf{C}}
\newcommand\body{{\cal B}}
\newcommand\Sone{\mathbf{S}^1}
\newcommand\Sthree{\mathbf{S}^3}
\newcommand\conf{\mathbf{q}}
\newcommand\xx{\mathbf{x}}
\newcommand\cost{C}
\newcommand\weight{W}
\newcommand\translation{\mathbf{t}}
\newcommand\tcolli{t_{coll\ i}}

@@

Some heuristics use random shortcuts on the initial guess combined with a 
trajectory re-building that returns ${C}^1$ shortcuts made of parabolas and lines 
(bang-bang control)~\cite{HauserFastSmooth}. These local refined trajectories 
are time-optimal since they comply with acceleration and velocity constraints.

@@

TO DESCRIBE:

~\cite{Guernane2011} (guiding shoots) is ``locally optimal'', not addressing high-DOF problems,

~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is using medial axis retractation for clearance and a partial random shortcut (applied only on random DOF) which can address the problem of ~\ref{decoupled_DOF_optimization} but has longer computation time and is not using the information returned by the checker of which links are colliding to guide the shortcut.

@@

\begin{figure}[b]
	\centering
	\includegraphics[width=8cm]{optim_grad.pdf}
	\caption{Illustration of one iteration of our optimization. $\xx_{i+1}$ 
	appears 
	to be in collision with the obstacle, the first colliding configuration 
	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The corresponding constraint will be computed in configuration $\xx_{i}(\tcolli)$.}
	\label{optim_grad}
\end{figure}

Let $T$ be a positive real number such that each path $\xx_i$ is a mapping from 
interval $[0,T]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(T) = q_{n+1}$ for all $i$. Let 
us denote by $\tcolli$ the abscissa of the first collision detected on path 
$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free (see 
Figure~\ref{optim_grad}). Thus in 
configuration $\xx_{i+1}(\tcolli)$ a collision has been 
detected. Two cases are possible:
\begin{enumerate}
\item the collision occurred between two bodies of the robot: $\body_1$ and $
\body_2$, or
\item the collision occurred between a body of the robot $\body_1$ and the 
environment.
\end{enumerate}
In the rest of this section, we will only consider the first case. Reasoning about the second case is similar, except that the 
constraint is on the transformation of $\body_1$ with respect to the environment.

\begin {figure}
  \centerline {
    \def\svgwidth {\linewidth}
    \graphicspath {{./images/}}
    \input {images/transformation-constraint.pdf_tex}
  }
  \caption {Transformation constraint notations.}
  \label{fig:transformation-constraint}
\end {figure}
Let $M^{*}\in SE(3)$ denote the homogeneous transformation between frames of $\body_1$  and $\body_2$ in configuration $\xx_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$. Let $R\in SO(3)$ (resp. $R^{*}$) and $\translation\in\real^3$ (resp. $\translation^{*}$) be the rotation and translation parts of $M$ (resp. $M^{*}$).
See also Figure~\ref{fig:transformation-constraint} for notations.

We denote by $F$ the mapping from $\CS^{wp}$ to $\real^6$ defined by
\begin{equation}\label{eq:constraint}
F (\xx) = \left(\begin{array}{c} R^T (\translation^{*} - \translation) \\
  \log (R^T R^{*})\end{array}\right)
\end{equation}
$\log (R^T R^{*})$ represents here a vector $\mathbf{v}$ of $\real^3$ such that
$R^T R^{*}$ is the rotation of axis $\mathbf{v}$ and of angle $\|\mathbf{v}\|$.
Note that $F$ vanishes for any path $\xx$ such that in configuration
$\xx(\tcolli)$ the relative position of $\body_2$ w.r.t. $\body_1$ is equal to $M^{*}$.
Let $r$ denotes the \textit{radius} of the bounding sphere of $\body_2$ centered at the origin of $\body_2$ local frame. Let us denote by $d$ the minimal distance between $\body_1$ and $\body_2$
in configuration $\xx_i(\tcolli)$. We have the following property.
\begin {property} \label {prop:collision free}
Let $\xx\in\CS^{wp}$ be a path and let us denote by $\translation$ and $\mathbf{v}$ respectively the 3 first coordinates and the 3 last coordinates of $F(\xx)$. If
$$
\|\translation\| + r \|\mathbf{v}\| \leq d
$$
then in configuration $\xx(\tcolli)$, $\body_1$ and $\body_2$ are collision-free.
\end {property}
The proof is straightforward: expressed in reference frame of $\body_1$, the motion of $\body_2$ between configurations $\xx(\tcolli)$ and $\xx_i(\tcolli)$ is a translation of vector $\translation$ followed by a rotation of angle $\|\mathbf{v}\|$. No point of $\body_2$ has moved by more than $\|\translation\| + r \|\mathbf{v}\|$. Therefore no collision is possible.

@@

We refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
programs (QP). This corresponds to \texttt{computeIterate} in Algorithm~\ref{algo:gradient}.

>>>>

% Math shortcuts :
\newcommand\real{\mathbb{R}}
\newcommand\p{\mathbf{p}}
\newcommand\pii{\mathbf{p_{i,i+1}}}
\newcommand\pij{\mathbf{p_{i,j}}}
\newcommand\gT{\tilde{g}^T}
\newcommand\g{\tilde{g}}
\newcommand\CS{\mathcal{C}}
\newcommand\dimCS{n_\mathbf{C}}
\newcommand\body{{\cal B}}
\newcommand\Sone{\mathbf{S}^1}
\newcommand\Sthree{\mathbf{S}^3}
\newcommand\conf{\mathbf{q}}
\newcommand\xx{\mathbf{x}}
\newcommand\cost{C}
\newcommand\weight{W}
\newcommand\translation{\mathbf{t}}
\newcommand\tcolli{t_{coll\ i}}
\newcommand\po{\mathbf{po}}

@@

Some heuristics use random shortcuts on the initial guess combined with a 
trajectory re-building that returns ${C}^1$ shortcuts made of parabolas and lines 
(bang-bang control)~\cite{HauserFastSmooth}. These local refined trajectories 
are time-optimal since they comply with acceleration and velocity constraints. ~\cite{Guernane2011}, also based on random shooting, is guiding configuration-generation with local holonimic considerations. But this method remains only locally optimal, and is not addressing high-DOF problems.
~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is using medial axis retractation for clearance and a partial random shortcut (applied only on random DOF) which can address the problem of ~\ref{decoupled_DOF_optimization} but has longer computation time and is not using the information returned by the checker of which links are colliding to guide which DOF to choose for the partial shortcut.

@@

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=9cm]{optim_grad.pdf}
	\caption{Illustration of one iteration of our optimization. $\xx_{i+1}$ 
	appears 
	to be in collision with the obstacle, the first colliding configuration 
	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The corresponding constraint will be computed in configuration $\xx_{i}(\tcolli)$.}
	\label{optim_grad}
\end{figure}

Let $T$ be a positive real number such that each path $\xx_i$ is a mapping from 
interval $[0,T]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(T) = q_{n+1}$ for all $i$. Let 
us denote by $\tcolli$ the abscissa of the first collision detected on path 
$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free (see 
Figure~\ref{optim_grad}). Thus in 
configuration $\xx_{i+1}(\tcolli)$ a collision has been 
detected. Two cases are possible:
\begin{enumerate}
\item the collision occurred between two bodies of the robot: $\body_1$ and $
\body_2$, or
\item the collision occurred between a body of the robot $\body_1$ and the 
environment.
\end{enumerate}
In the rest of this section, we will only consider the first case. Reasoning 
about the second case is similar, except that the 
constraint is on the position of $\body_1$ with respect to the environment.

\begin{figure}[b]
	\centering
	\includegraphics[width=15cm]{contact-points.pdf}
	\caption{Constraint definition based on contact points.}
	\label{contact-points}
\end{figure}

The principle of the method is to compute a constraint on the backtracked free-
configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
(\tcolli)$. To handle this, we introduce a one-dimensional constraint based on the 
orthogonal direction of the encountered collision.

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be the 
contact point expressed in the global frame ${\cal R}_0$ of each of the two 
bodies. We compute (see Figure~\ref{contact-points}) the associated points $\po_1$ 
and $\po_2$ in the free-configuration $\xx_{i}(\tcolli)$, and thus 
$\mathbf{u}=\po_2-\po_1$.

Let $\translation^{*}\in \real^3$ denotes the position between frames of $\body_1$  
and $\body_2$ in configuration $\xx_i(\tcolli)$. Let $\translation \in 
\real^3$ (resp. $R\in SO(3)$) denote the same position (resp. rotation) in 
configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$.

We denote by $F$ the mapping from $\CS^{wp}$ to $\real$ defined by
\begin{equation}\label{eq:constraint}
F (\xx) =  \mathbf{u}^T \, R^T (\translation^{*} - \translation)
\end{equation}
Note that $F$ vanishes for any path $\xx$ such that in configuration
$\xx(\tcolli)$ the relative position of $\body_2$ w.r.t. $\body_1$ is equal to $
\translation^{*}$ or if the relative position $\translation^{*} - \translation$ is 
orthogonal to $\mathbf{u}$.

@@

\vspace{0.2cm}

Let $q_{k\,i}$ denotes waypoint $k$ along path $x_i$.
The constraint on the position at $t_{coll\,i}$ is a function that depends only
on two waypoints $q_{k\,i}$, $q_{k+1\,i}$, with $0\leq k\leq n$. There exists 
$\beta\in[0,1]$ such that for any $j$:
$$
\xx_j (t_{coll\,i}) = (1-\beta) \conf_{k\,j} + \beta \conf_{k+1\,j}
$$
Therefore, the Jacobian $J_F$ of constraint~(\ref{eq:constraint}) with respect to 
$x$ is
$$
J_F = \left(\begin{array}{cccccc}
0 & \cdots & (1-\beta)\frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) &
\beta\frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \cdots & 0
\end{array}
\right)
$$
Both $F$ and $\frac{\partial F}{\partial q}$ are provided by our path planning framework~\cite{hpp} as differentiable functions.

\vspace{0.2cm}

Each time a new constraint is added to the resolution, we stack its jacobian $J_F$ to the previous ones. This corresponds to \texttt{addCollisionConstraints} in 
Algorithm~\ref{algo:gradient}. Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic programs (QP) (function \texttt{computeIterate} in 
Algorithm~\ref{algo:gradient}).
